---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# recPhyloParse

<!-- badges: start -->
  [![R-CMD-check](https://github.com/fmarotta/recPhyloParse/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/fmarotta/recPhyloParse/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/fmarotta/recPhyloParse/branch/main/graph/badge.svg)](https://app.codecov.io/gh/fmarotta/recPhyloParse?branch=main)
<!-- badges: end -->

recPhyloParse provides a way to import reconciliated phylogenies into R in a way that makes it easy to plot the trees with ggplot2. The starting point is an XML file following the [recPhyloXML format](http://phylariane.univ-lyon1.fr/recphyloxml/), which will be parsed into an [R6 object](https://r6.r-lib.org/index.html). During the parsing, the coordinates and attributes of all the plot elements will be appropriately extracted into data.frames, so that standard ggplot2 geoms can be used.

## Installation

You can install the development version of recPhyloParse from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("fmarotta/recPhyloParse")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r}

library(recPhyloParse)
library(ggplot2)

recphylo_file <- system.file("extdata", "example_1.recphyloxml", package = "recPhyloParse")
recphylo <- RecPhylo$new(recphylo_file, branch_length_scale = 5, x_padding = 4)

ggplot() +
  geom_line(data = recphylo$spEdges, aes(x, y, group = group)) +
  geom_line(data = recphylo$recGeneEdges, aes(x, y, group = group, color = gsub("l+$", "", lineage), linetype = event_type), show.legend = F) +
  scale_linetype_manual(values = c("loss_v" = 2, "transferBack" = 3), na.value = 1) +
  theme_void() +
  NULL

```

## Features

- Parse and draw recPhyloXML trees with ggplot2
- Three linear layouts: real branch lengths, uniform branch lengths, aligned leaves
- Corresponding circular layouts "for free" thanks to coord_polar()
- Swap left and right children of a species node
- Swap children of a duplication event
- Import branch lengths from another species tree file
- Add custom annotation for species nodes, gene nodes, species edges, gene edges

## Bugs and missing features

- We support only recphyloxml files with exactly one gene and one species tree
- BifurcationOut events (as well as free-living species) are currently not supported
- We don't have the ability to infer species tree from gene tree (this is a prerequisite for implementing the previous two, maybe I'll give it a shot)
- We make no attempt to minimise the edge crossings for lateral
  transfers (though you can manually swap left/right children to make the tree look better)
- We don't automatically scale the plot to make it look nicer (planned)
- We do not add clade attributes and internal elements as columns in the
  resulting data.frame (we could add all these:
  <http://www.phyloxml.org/documentation/version_1.20/phyloxml.html>)
